MLH1 Data Report

### 2019-09-17 doc/Main_Report.rmd
some updates to main report doc

Cell images are quantified in batches after file names are anonymized. The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified. A quality score is given; 1 to 5 (best to worst).

all cell image file names are anonymized and moved into batches of a few hundred images.

@@

@@

@@

Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "female",], 
                 .(category,subsp),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci)
)

tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "female",], 
                 .(category,subsp),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci)
)

@@

Ranges for male means across subspecies. Ranges of means (not ranges of all).

(try turning this into a table...)
Dom            female Dom
old,   new      old     new
22-24  23-25    24-27  25-29

Musc           female Musc
23-31   25-31   25-27   26-29
22-25   22-27    25       25
(ASTmale has 27 mean, outlier)

Outgroup       female outgroup
25-26   26-27    27    29-31

I think the ranges across the groups

### Assesment of Quality Scores

@@

#are there no KAZ cells of high quality?
qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey" ))
high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
```

@@

lm.me  <- lm(male.data$mean_co ~ male.data$age.weeks+male.data$strain.x)

#summary(lm.me)
```

@@

@@

Rough ideas for models of evolution of Heterochiasmy and current idea for how to test them.

@@

run analysis for whole SC skel measures

@@

looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Musc",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
#very possible that the dimesions and pixel density are skrewing with these values
# I am suspcious that there is such a wide range for the female data

looky.DOM <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))

>>>>

Cell images are quantified in batches after file names are anonymized. The number of MLH1 foci, number of "achiasmate" and asynasped bivalents are quantified. A quality score is given; 1 to 5 (best to worst). All cell image file names are anonymized and moved into batches of a few hundred images.

@@

#in the chrm proportions data is another set for eastimating error

@@

*dominant effect* (F1s)

@@

Q12_MLH1_means_males <- ddply(.data=Q12_MLH1_DF[Q12_MLH1_DF$sex == "male",], 
                 .(category,subsp),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci)
)

tot.DF_MLH1_means_males <- ddply(.data=MLH1_data[MLH1_data$sex == "male",], 
                 .(category,subsp),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean(adj_nMLH1.foci)
)

@@

Ranges of category means across subspecies.

| Type | male Dom | female Dom | | male Musc low | male Musc high | Musc female | | male Outgroup | female outgroup |
| --- | --- | --- | --- | --- | --- | --- | --- | --- | --- |
| Total cells | 22-24 |  24-27 |  | 23-31 | 25-31 | 26-29 | | 25-26 |27 |
| High quality cells | 23-25 | 25-29 | | 22-25 | 22-27 | 25 | | 26-27 | 29-31 |
(ASTmale in Musc high, has a 27 mean which might be an outlier)
I think the ranges across the groups is generally consistant.

(try looking at <5 quality scores?)

@@

#are there no KAZ cells of high quality?
qq <- ggplot(highest_qual_df, aes(x = mouse, y = adj_nMLH1.foci)) + geom_boxplot(data = highest_qual_df, aes(fill = factor(strain) ) ) +  ggtitle("Cells with quality scores of 2 or 1") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue",  "coral1", "#E69F00", "red","yellowgreen", "green","purple", "black", "grey", 
             "grey2","grey3","grey1","grey","grey","grey"))
high_qual_boxplot <- qq + facet_wrap(~ sex, scales="free") #this is breaking the kniting... 
#high_qual_boxplot
#these boxplots look bad
```

Higher quality scored cells have more MLH1 foci. (some of the ranges get smaller with higher quality cells)

(was I referring to normale distributions?)
means all shift up, what about variance, range, lowest cut off

> *Do the directions of the differences across groups change due to quality scores?*

(polymorphism vs divergence patterns)

@@

male.data.dom <- male.data[male.data$subsp.y == "Dom",]

male.data.musc.low <- male.data[(male.data$strain.x == "KAZ" | male.data$strain.x == "TOM" | male.data$strain.x == "AST"),]

male.data.musc.high <- male.data[(male.data$strain.x == "PWD" | male.data$strain.x == "MSM" | male.data$strain.x == "MOLF" | male.data$strain.x == "SKIVE"),]

male.data.outgroup <- male.data[(male.data$subsp.y == "Spretus" | male.data$subsp.y == "Spic" | male.data$subsp.y == "Caroli"),]

lm.me.dom  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.dom)

lm.me.musc.low  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.low)

lm.me.musc.high  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.musc.high)
#significant differences across strains in musc high

lm.me.OG  <- lm(mean_co ~ age.weeks+strain.x, data = male.data.outgroup)

summary(lm.me.OG)
#for all models, mostly it was that the intercept was significant, and mot much else
```

```{r female.age.effect, echo=FALSE}
#female age effect

#female.meta.data <- meta.data[meta.data$sex == "female",]
#how much data to I have to work with?
#there's a fair number of rows without emb or neonate
#I could probably assume all of these are neonates

#length( !is.na(female.meta.data$female.stage) )

#female.meta.data <- female.meta.data[female.meta.data$female.stage]

#how much formating do I have to do?
#can't use AP_mouse_table since that is summarized
#meta.data
#grepl notes, "neo"

#some have e-number for embryonic
#I might have to

I will try doing a logistic regression for embryonic vs neonate. The embyonic estimates likely aren't very accurate, and there won't be good sampling across time points.

@@

(in progress)

@@

Ideas for models of evolution of Heterochiasmy and current idea for how to test them.

@@

data has been collection for RWang's total SC. These results are consistant with the patterns seen in the bivalent level resutls. (total SC PWD > KAZ )

@@

looky <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Musc",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
#very possible that the dimesions and pixel density are skrewing with these values
# I am suspcious that there is such a wide range for the female data
looky

looky.DOM <- ggplot(data = mlh1.skel.og[mlh1.skel.og$subsp == "Dom",], aes(y=skel_size, x=nMLH1.foci, color=mouse) ) + geom_jitter() + facet_wrap(sex~strain)+theme(legend.position="none")  +ylim(c(0,2000))
looky.DOM
